The effects of multi environmental exposure on Tuberculosis Incidence in Asian and African countris

### 2025-02-02 products/manuscript/Manuscript.qmd
Install renv() to make the project reproducible
fixed some typo

---
title: "The effects of multi environmental exposure on Tuberculosis Incidence in Asian and African countris"
author: ""
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
---

@@

* Muhammad Nasir$^{1,2,*}$ (ORCID: 0000-0000-1234-5678) 
* Second Author$^{3, *}$
* Third Autor$^{3, \dagger}$
* Last Author$^{1, 2, \land}$

@@

1. College of Public Health, University of Georgia, Athens, GA, USA.
2. Another Affiliation, Athens, GA, USA.
3. Yet another Affiliation.

@@

$\land$ Corresponding author: some@email.com

@@

# Summary/Abstract
_Write a summary of your project._

@@

Tuberculosis is old infectious disease which still exists and become major public health in the world, particularly in developing countries. Tuberculosis is linked to environmental factors. This study will analyze the effects of multi environmental factors on tuberculosis incidence in Asia and Afirca regions, where the majority of tuberculosis coming from these regions. The environmental factors and country economic and hman development index as covariates.

@@

The study will use multi sources of data including World Bank data sources, World Health Organization (WHO), United Nation for Children Fund (Unicef), and UN-HABITAT, Emissions Database for Global Atmospheric Researcg (EDGAR), and Institute for Health Metrics and Evaluation (IHME). The data contains informations of all variables in all countries globally from 2000 to 2020. The variables will include  proportion of urban opulation living in slum areas, percentage of popuation using safely managed sanitation services, percentage of population using at least basic drinking water sevice, fule for cooking, Population-weighted exposure to ambient PM2.5 pollution, total green houses, and will include GDP/ country economic status, human development index, and Universal Health Caverage as covariates.

Note: I have downloaded the dataset, but I have not conducted any cleaning or exploration on the dataset. 
List of data sourses and website is provided in an exel fine in my data folder. I will include them in my reference for the next steps.

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._

@@

RQ: What enironmental factors asociated with tuberculosis incidence?

H0: there is no relationship between proportion ofurban population living in slum areas and tuberculosis. 
H0: thee is no relationship between safely managed santation and tuberculosis incidence 
H0: there is no relatinship between drinking water and tuberculosis incidence 
H0: there is no relationship between fule for cooking and tuberculosis incidence 
H0: there is no relationship between PM 2.5 exposure and tuberclosis incidence 
H0: there is no association between total green gass houses and tuberculosis incidence

Ha: there is  relationship between proportion ofurban population living in slum areas and tuberculosis. 
Ha: thee is relationship between safely managed santation and tuberculosis incidence 
Ha: there is relatinship between drinking water and tuberculosis incidence 
Ha: there isrelationship between fule for cooking and tuberculosis incidence 
Ha: there is relationship between PM 2.5 exposure and tuberclosis incidence 
Ha: there is association between total green gass houses and tuberculosis incidence

@@

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

I haven't decide about data analysis methods. However, I am thinking to perform time series analysis and multi linear regression (if suitable) or Generealized linear model. I am still in the process of reading more papers to get more informatin which methods best suit this data and answer the research questions.

@@

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

I will add the code.

## Statistical analysis
_Explain anything related to your statistical analyses._

@@

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@@

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@@

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

@@

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

>>>>

---
title: "The effects of multi environmental exposure on Tuberculosis Incidence in Asian and African countris"
author: ""
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
editor: 
  markdown: 
    wrap: 72
---

@@

-   Muhammad Nasir$^{1,2,*}$ (ORCID: 0000-0000-1234-5678)
-   Second Author$^{3, *}$
-   Third Author$^{3, \dagger}$
-   Last Author$^{1, 2, \land}$

@@

1.  College of Public Health, University of Georgia, Athens, GA, USA.
2.  Faculty of Public Health, Universitas Tadulako, Palu, Indonesia.
3.  Yet another Affiliation.

@@

$\land$ Corresponding author: some\@email.com

@@

# Summary/Abstract

*Write a summary of your project.*

@@

Tuberculosis is old infectious disease which still exists and become
major public health in the world, particularly in developing countries.
Tuberculosis is linked to environmental factors. This study will analyze
the effects of multi environmental factors on tuberculosis incidence in
Asia and Africa regions, where the majority of tuberculosis coming from
these regions. The environmental factors and country economic and human
development index as covariates.

@@

The study will use multi sources of data including World Bank, World
Health Organization (WHO), United Nation for Children Fund (Unicef),
UN-HABITAT, Emissions Database for Global Atmospheric Research (EDGAR),
and Institute for Health Metrics and Evaluation (IHME). Different
dataset contains different variable, however all dataset have same year,
country name and country codes that can be used to combine the dataset.
The data also contain information of countries globally from 2000 to
2020. The variables will include proportion of urban population living
in slum areas, percentage of population using safely managed sanitation
services, percentage of population using at least basic drinking water
service, fuel for cooking, Population-weighted exposure to ambient PM2.5
pollution, total green houses, and will include GDP/ country economic
status, human development index, and Universal Health Coverage as
covariates.

Note: I have downloaded the dataset, but I have not conducted any
cleaning or exploration on the dataset. List of data sources and website
is provided in an excel fine in my data folder. I will include them in
my reference for the next steps.

## Questions/Hypotheses to be addressed

*State the research questions you plan to answer with this analysis.*

@@

RQ: What environmental factors associated with tuberculosis incidence?

H0: there is no relationship between proportion of urban population
living in slum areas and tuberculosis.

H0: thee is no relationship between safely managed sanitation and
tuberculosis incidence

H0: there is no relationship between drinking water and tuberculosis
incidence

H0: there is no relationship between fuel for cooking and tuberculosis
incidence

H0: there is no relationship between PM 2.5 exposure and tuberculosis
incidence

H0: there is no association between total green gas houses and
tuberculosis incidence

Ha: there is a relationship between proportion of urban population
living in slum areas and tuberculosis.

Ha: there is a relationship between safely managed sanitation and
tuberculosis incidence

Ha: there is a relationship between drinking water and tuberculosis
incidence

Ha: there is a relationship between fuel for cooking and tuberculosis
incidence

Ha: there is a relationship between PM 2.5 exposure and tuberculosis
incidence

Ha: there is a association between total green gas houses and
tuberculosis incidence

@@

*Describe your methods. That should describe the data, the cleaning
processes, and the analysis approaches. You might want to provide a
shorter description here and all the details in the supplement.*

I haven't decide about data analysis methods. However, I am thinking to
perform time series analysis and multi linear regression (if suitable)
or Gene-realized linear model (we will see after conducting data
exploration). I am still in the process of reading more papers to get
more information which methods best suit this data and answer the
research questions.

@@

## Data aquisition

Most of the data is downloaded from Worlbank Dataset. However, those
data are originally from many different courses/ organizations, such as
World Bank, WHA, UNICEF, UN-HABITAT, EDGAR, and IHME.

## Data import and cleaning

*Write code that reads in the file and cleans it so it's ready for
analysis. Since this will be fairly long code for most dataset, it might
be a good idea to have it in one or several R scripts. If that is the
case, explain here briefly what kind of cleaning/processing you do, and
provide more details and well documented code somewhere (e.g. as
supplement in a paper). All materials, including files that contain
code, should be commented well so everyone can follow along.*

Dataset are either in excel or CSV files. Several steps from loading to
data analysis:

-   to load the data set in this part using read_excel or read_csv.

-   to combine them base on country name/ code and years.

<!-- -->

-   to drop unwanted variables

-   to select Asian and African Countries only

-   to drop N/A or missing value

-   to conduct data exploration

-   to decide suitable analysis method

-   to perform data analysis

## Statistical analysis

*Explain anything related to your statistical analyses.*

@@

*Use a combination of text/tables/figures to explore and describe your
data. Show the most important descriptive results here. Additional ones
should go in the supplement. Even more can be in the R and Quarto files
that are part of your project.*

@@

*To get some further insight into your data, if reasonable you could
compute simple statistics (e.g. simple models with 1 predictor) to look
for associations between your outcome(s) and each individual predictor
variable. Though note that unless you pre-specified the outcome and main
exposure, any "p\<0.05 means statistical significance" interpretation is
not valid.*

@@

*Use one or several suitable statistical/machine learning methods to
analyze your data and to produce meaningful figures, tables, etc. This
might again be code that is best placed in one or several separate R
scripts that need to be well documented. You want the code to produce
figures and data ready for display as tables, and save those. Then you
load them here.*

@@

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your
analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references
will automatically be placed at the end*
